Parameter estimation

### 2023-04-22 SFE-article/Sections/Model.tex
21.04

	\subsubsection{Continuity equation} \label{CH: Continuity}
	The continuity equation for the fluid phase is derived in Appendix \ref{CH: Gouverning equations}. When the cross-sectional area of the channel $A_f$ is specified as a function of the void fraction ${\color{magenta}\phi}(z)$ (where ${\color{magenta}\phi}$ is the void fraction of the bed and $A$ is the cross-section of the empty extractor), the continuity equation takes the form: \todo[]{assign color to v and u}

@@

	Analogously, the superficial velocity mught be introduced to the model and defined as

@@

	Here, ${\color{blue}c_f}(t,z), {\color{blue}c_s}(t,z),$ and ${\color{blue}T}(t,z)$ represent the concentration of solute in the fluid phase, concentration of solute in the solid phase, and temperature, respectively. ${\color{blue}r_e}(t,z)$ is a mass transfer kinetic term. ${\color{red}F}(t)$ is the mass flow rate, ${\color{red}P}(t)$ is the pressure, ${\color{magenta}\epsilon}$ is the void fraction of the bed, ${\color{orange}\rho_f}({\color{blue}T}(t,z),{\color{red}P}(t))$ is the fluid density, ${\color{magenta}\rho_s}$ is the solids density, ${\color{orange}D^M_e}({\color{blue}T}(t,z),{\color{red}P}(t),{\color{red}F}(t))$ is the axial mass diffusion coefficient, and $u$ is the superficial velocity.

@@

	The kinetic term in this study is based on the two-film theory proposed by \citet{Reverchon1996} and the mass transfer kinetic is given by equation \ref{Model_kinetic_basic}. This equation takes into account the overall diffusion coefficient and the concentration gradient, which acts as the driving force for the process.

@@

	As the solvent flows through the bed, $CO_2$ molecules diffuse into the pores and adsorb on the particle surface to form an external fluid film around the solid particles due to the solvent-solid matrix interactions. The effect of Knudsen diffusion is negligible in this process, as the mean free path of the molecule is much smaller than the pore diameter. The dissolved solute diffuses from the particle's core through the solid-fluid interface, the pore, and the film into the bulk. Figure \ref{fig: SFE_Mechanism} illustrates the mass transfer mechanism, where the mean solute concentration in the solid phase is denoted as ${\color{blue}c_s}$ and the equilibrium concentrations at the solid-fluid interface are denoted as ${\color{blue}c_s^*}$ and ${\color{blue}c_P^*}$, respectively, for solid and fluid phases. The concentration of the solutes in the fluid phase in the centre of the pore is denoted as ${\color{blue}c_{P}}$. As the solute diffuses through the pore, its concentration changes and reaches ${\color{blue}c_{Pf}}$ at the opening of the pore. The solute then diffuses through the film around the particle and reaches a concentration in the bulk ${\color{blue}c_f}$. The two-film theory describes the solid-fluid interface inside the pore. The overall mass transfer coefficient can be determined if the relation between the solute concentration in one phase and its equilibrium concentration is known.

@@

		\citet{Bulley1984} suggest a process where the driving force for extraction is given by the difference between concentration of the solute in the bulk, ${\color{blue}c_f}$, and in the centre of the pore, ${\color{blue}c_P^*}$. The concentration ${\color{blue}c_P^*}$ is in equilibrium with ${\color{blue}c_s}$ according to an equilibrium relationship. The rate of extraction is thus ${\color{blue}r_\text{e}}\left({\color{blue}c_f} - {\color{blue}c^*_P}({\color{blue}c_s})\right)$.

@@

		According to \citet{Bulley1984}, a linear equilibrium relationship (equation  \ref{Linear_equilibirum}) can be used to find an equilibrium concentration of the solute in the fluid phase ${\color{blue}c_f^*}(t,z)$ is based on concentration of the solute in the solid phase ${\color{blue}c_s}(t,z)$

@@

			Following \citet{Elliott2011} or \citet{Gmehling2019}, a real gases internal energy definition can be obtained from the departure functions:

@@

			If a gas is consider to be perfectly caloric (${\color{blue}e}(t,z)=C_v {\color{blue}T}(t,z)$), then the energy equation can be written in form of temperature. The perfectly caloric gas can be seen as the special case of a real gas, where the second term of the equation \ref{EQ: internal_energy_def} goes to zero.

			For real gases it is complicated to write the heat balance in terms of temperature, but it be can used directly in the form of internal energy, as it is given by equation \ref{EQ: CompressibleEuler_3}. In such a case, the temperature (which is used as the input to some other functions) needs to be recover from the internal energy. A relation for the internal energy can be obtained from an equation of state. For Peng-Robinson, such a relation is given by equation \ref{EQ: internal_energy_PR} as presented by \citet{Elliott2011}.

@@

			To solve equation \ref{EQ: internal_energy_PR}, values of temperature, pressure and density needs to be know. If an equation of state is introduce, then only two out three variables needs to be know as the third one can be calculated, this can be represented as follow

@@

			If the value of internal energy ${\color{blue}e}(t,z)$ is know from time evolution of the energy equation (\ref{EQ: CompressibleEuler_3}), and pressure is know from measurement, then the temperature can be reconstructed. A roootfinder can be used to find a value of temperature, which minimize the difference between value of internal energy coming from the time evolution and the output from equation \ref{EQ: internal_energy_PR}. Such a procedure allow to find local temperature along spatial direction ${\color{blue}z}$ and needs to be repeated every time-step.

@@

			The main advantage of this formulation is the presence of term $\partial {\color{red}P}(t) / \partial t $, which allow to directly affect the system through the change of thermodynamic pressure (which is a control variable).\todo{ "For high pressures, the influence of the intermolecular forces on the enthalpy has to be taken into account, usually by applying a cubic equation of state. In most cases, these forces are attractive, so that additional energy is necessary to move the molecules away from each other, that is, to lower the density. If this energy is not added, the substance cools down when it is expanded" (\citet{Gmehling2019}). So, pressure reduction cause gas expansion and temperature drop.}
			If an equation of state is known, the temperature can to be recovered from the enthalpy. The enthalpy is related to the pressure and temperature through the following equation:

@@

			If the value of enthalpy is know from the time evolution and pressure can be measured, then the equation \ref{EQ:Enthalpy_root} can be solved for temperature to recover the temperature profile. By applying the departure functions to Peng-Robinson equation of state, the relation \ref{EQ:Enthalpy_root} can be expressed directly through equation \ref{EQ:Enthalpy_PR} as presented in \ref{CH:Enthalpy} or given by \citet{Gmehling2019}.

@@

		\todo[inline]{Add short section about boundary conditions and comment on the used numerical methods}

@@

		The efficiency of the process (the yield) is calculated according to equations \ref{Model_measurment_1} to \ref{Model_measurment_2}, which evaluate the mass of solute at the exit of the extraction unit and sums it. The integral form of the measurement equation can be transformed into the differential form, and augmented with model equations to be solved simultaneously.

@@

		\iffalse

		\subsubsection{State-space representation} \label{CH: State_space}
		It is assumed that the solvent is free of solute at the entrance of the extractor and that all the solid particles have the same initial solute content ${\color{magenta}q_0}$. Moreover, it is considered that the initial temperature of the extractor in every place is equal to ${\color{magenta}T_0}$. Therefore, the initial conditions employed in the simulation are:

		{\footnotesize
			\begin{subequations}
				\begin{align*}
					{\color{blue}c}(t = 0, z) &= 0   \\
					{\color{blue}q}(t = 0, z) &= {\color{magenta}q_0} \\
					{\color{blue}T}(t = 0, z) &= {\color{magenta}T_0}
				\end{align*}
		\end{subequations} }

@@

		{\footnotesize
			\begin{align}
				\begin{bmatrix}
					\cfrac{\partial {\color{blue}c}(t,z)}{\partial t}\\
					\cfrac{\partial {\color{blue}q}(t,z)}{\partial t}\\
					\cfrac{\partial {\color{blue}T}(t,z)}{\partial t} 
				\end{bmatrix}
				& =
				\begin{bmatrix}
					{\color{blue}\phi_1} \left( {\color{blue}c}(t,z),{\color{blue}q}(t,z),{\color{blue}T}(t,z); {\color{magenta}\theta} \right)\\
					{\color{blue}\phi_2} \left( {\color{blue}c}(t,z),{\color{blue}q}(t,z),{\color{blue}T}(t,z); {\color{magenta}\theta} \right)\\
					{\color{blue}\phi_3} \left( {\color{blue}c}(t,z),{\color{blue}q}(t,z),{\color{blue}T}(t,z); {\color{magenta}\theta} \right)
				\end{bmatrix} = {\color{blue}\phi} \left( t,z; {\color{magenta}\theta} \right) = \cfrac{\partial {\color{blue}\chi}(t,z)}{\partial t}
		\end{align} }

@@

		The method of lines is used to transform the process model equations into a set of ODEs denoted as ${\color{blue}G}({\color{blue}x}(t);{\color{magenta}p})$. The partial derivatives in $z$-direction are computed using a first-order and second-order finite difference approximation. The backward finite difference is used to approximate first-order derivative, while the central difference scheme is used to approximate second-order derivative. The length of the fixed bed is divided into $N_z$ equally distributed points in $z$-direction. Each function ${\color{blue}\phi_i}$ is transformed to a corresponding set of $N_z$ discretized equations denoted as ${\color{blue}G}_{i\times N_z+1}$ to ${\color{blue}G}_{(i+1)\times N_z}$, where $i$ corresponds to the process model equation. The state-space model ${\color{blue}\chi}(t,z)$ after the discretization is represented by $\dot{{\color{blue}x}}(t)$ (equation  \ref{discretization}).

			{\footnotesize
				\begin{align*} \label{discretization}
					\dot{{\color{blue}x}}(t) &= \cfrac{d {\color{blue}x}(t)}{d t} = 
					\begin{bmatrix}
						\cfrac{d {\color{blue}c_{f,1}}(t)}{d t} 	  \\
						\vdots					  \\
						\cfrac{d {\color{blue}c_{f,N_z}}(t)}{d t} \\
						\\ \hline \\
						\cfrac{d {\color{blue}c_{s,1}}(t)}{d t} 	  \\
						\vdots					  \\
						\cfrac{d {\color{blue}c_{s,N_z}}(t)}{d t} \\
						\\ \hline \\
						\cfrac{d {\color{blue}T_1}(t)}{d t} 	  \\
						\vdots 					  \\
						\cfrac{d {\color{blue}T_{N_z}}(t)}{d t}
					\end{bmatrix}
					=
					\underbrace{\begin{bmatrix}
							{\color{blue}G_1} \left( {\color{blue}x}(t),{\color{blue}q}(t),{\color{blue}T}(t); {\color{magenta}p} \right)\\ 
							\vdots\\ 
							{\color{blue}G_{N_z}} \left( {\color{blue}c}(t),{\color{blue}q}(t),{\color{blue}T}(t); {\color{magenta}p} \right)\\ 
							\\ \hline \\ \\
							{\color{blue}G_{N_z+1}} \left( {\color{blue}c}(t),{\color{blue}q}(t),{\color{blue}T}(t); {\color{magenta}p} \right)\\ 
							\vdots\\
							{\color{blue}G_{2N_z}} \left( {\color{blue}c}(t),{\color{blue}q}(t),{\color{blue}T}(t); {\color{magenta}p} \right)\\ 
							\\ \\ \hline \\ 
							{\color{blue}G_{2N_z+1}} \left( {\color{blue}c}(t),{\color{blue}q}(t),{\color{blue}T}(t); {\color{magenta}p} \right) \\
							\vdots\\
							{\color{blue}G_{3N_z}} \left( {\color{blue}c}(t),{\color{blue}q}(t),{\color{blue}T}(t); {\color{magenta}p} \right)
					\end{bmatrix}}_{{\color{blue}G} \left( {\color{blue}x}(t); {\color{magenta}p} \right)} 
			\end{align*} }

			where ${\color{blue}x} \in \mathbb{R}^{N_x = 3N_z} $ and ${\color{magenta}p} \in \mathbb{R}^{N_p =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of model parameters, $N_{u}$ is the number of control variables.

@@

			\fi

>>>>

	\subsubsection{Continuity equation} \label{CH: Continuity}
	The continuity equation for the fluid phase is derived in Appendix \ref{CH: Gouverning equations}. When the cross-sectional area of the channel $A_f$ is specified as a function of the void fraction ${\color{magenta}\phi}(z)$ (where ${\color{magenta}\phi}$ is the void fraction of the bed and $A$ is the cross-section of the empty extractor), the continuity equation takes the form: \todo[]{assign a color to v and u}

@@

	Analogously, the superficial velocity might be introduced to the model and defined as

@@

	Here, ${\color{blue}c_f}(t,z), {\color{blue}c_s}(t,z),$ and ${\color{blue}T}(t,z)$ represent the concentration of solute in the fluid phase, the concentration of solute in the solid phase, and temperature, respectively. ${\color{blue}r_e}(t,z)$ is a mass transfer kinetic term. ${\color{red}F}(t)$ is the mass flow rate, ${\color{red}P}(t)$ is the pressure, ${\color{magenta}\epsilon}$ is the void fraction of the bed, ${\color{orange}\rho_f}({\color{blue}T}(t,z),{\color{red}P}(t))$ is the fluid density, ${\color{magenta}\rho_s}$ is the solids density, ${\color{orange}D^M_e}({\color{blue}T}(t,z),{\color{red}P}(t),{\color{red}F}(t))$ is the axial mass diffusion coefficient, and $u$ is the superficial velocity.

@@

	The kinetic term in this study is based on the two-film theory proposed by \citet{Reverchon1996}, and the mass transfer kinetic is given by equation \ref{Model_kinetic_basic}. This equation takes into account the overall diffusion coefficient and the concentration gradient, which acts as the driving force for the process.

@@

	As the solvent flows through the bed, $CO_2$ molecules diffuse into the pores and adsorb on the particle surface to form an external fluid film around the solid particles due to the solvent-solid matrix interactions. The effect of Knudsen diffusion is negligible in this process, as the mean free path of the molecule is much smaller than the pore diameter. The dissolved solute diffuses from the particle's core through the solid-fluid interface, the pore, and the film into the bulk. Figure \ref{fig: SFE_Mechanism} illustrates the mass transfer mechanism, where the mean solute concentration in the solid phase is denoted as ${\color{blue}c_s}$ and the equilibrium concentrations at the solid-fluid interface are denoted as ${\color{blue}c_s^*}$ and ${\color{blue}c_P^*}$, respectively, for solid and fluid phases. The concentration of the solutes in the fluid phase in the center of the pore is denoted as ${\color{blue}c_{P}}$. As the solute diffuses through the pore, its concentration changes and reaches ${\color{blue}c_{Pf}}$ at the opening of the pore. The solute then diffuses through the film around the particle and reaches a concentration in bulk ${\color{blue}c_f}$. The two-film theory describes the solid-fluid interface inside the pore. The overall mass transfer coefficient can be determined if the relationship between the solute concentration in one phase and its equilibrium concentration is known.

@@

		\citet{Bulley1984} suggests a process where the driving force for extraction is given by the difference between the concentration of the solute in bulk, ${\color{blue}c_f}$, and in the center of the pore, ${\color{blue}c_P^*}$. The concentration ${\color{blue}c_P^*}$ is in equilibrium with ${\color{blue}c_s}$ according to an equilibrium relationship. The rate of extraction is thus ${\color{blue}r_\text{e}}\left({\color{blue}c_f} - {\color{blue}c^*_P}({\color{blue}c_s})\right)$.

@@

		According to \citet{Bulley1984}, a linear equilibrium relationship (equation  \ref{Linear_equilibirum}) can be used to find an equilibrium concentration of the solute in the fluid phase ${\color{blue}c_f^*}(t,z)$ is based on the concentration of the solute in the solid phase ${\color{blue}c_s}(t,z)$

@@

			Following \citet{Elliott2011} or \citet{Gmehling2019}, a real gas internal energy definition can be obtained from the departure functions:

@@

			If a gas is considered to be perfectly caloric (${\color{blue}e}(t,z)=C_v {\color{blue}T}(t,z)$), then the energy equation can be written in form of temperature. The perfectly caloric gas can be seen as the special case of a real gas, where the second term of the equation \ref{EQ: internal_energy_def} goes to zero.

			For real gases, it is complicated to write the heat balance in terms of temperature, but it be can be used directly in the form of internal energy, as it is given by equation \ref{EQ: CompressibleEuler_3}. In such a case, the temperature (which is used as the input to some other functions) needs to be recovered from the internal energy. A relation for the internal energy can be obtained from an equation of state. For Peng-Robinson, such a relation is given by equation \ref{EQ: internal_energy_PR} as presented by \citet{Elliott2011}.

@@

			To solve equation \ref{EQ: internal_energy_PR}, values of temperature, pressure, and density need to be known. If an equation of state is introduced, then only two out of three variables need to be obtained as the third one can be calculated, this can be represented as follow

@@

			If the value of internal energy ${\color{blue}e}(t,z)$ is known from the time evolution of the energy equation (\ref{EQ: CompressibleEuler_3}), and pressure is known from measurement, then the temperature can be reconstructed. A roootfinder can be used to find a value of temperature, which minimizes the difference between the value of internal energy coming from the time evolution and the output from equation \ref{EQ: internal_energy_PR}. Such a procedure allow to find local temperature along spatial direction ${\color{blue}z}$ and needs to be repeated every time-step.

@@

			The main advantage of this formulation is the presence of term $\partial {\color{red}P}(t) / \partial t $, which allows it to directly affect the system through the change of thermodynamic pressure (which is a control variable).\todo{ "For high pressures, the influence of the intermolecular forces on the enthalpy has to be taken into account, usually by applying a cubic equation of state. In most cases, these forces are attractive, so additional energy is necessary to move the molecules away from each other, that is, to lower the density. If this energy is not added, the substance cools down when it is expanded" (\citet{Gmehling2019}). So, pressure reduction cause gas expansion and temperature drop.}
			If an equation of state is known, the temperature can to be recovered from the enthalpy. The enthalpy is related to the pressure and temperature through the following equation:

@@

			If the value of enthalpy is known from the time evolution and pressure can be measured, then the equation \ref{EQ:Enthalpy_root} can be solved for the temperature to recover the temperature profile. By applying the departure functions to Peng-Robinson equation of state, the relation \ref{EQ:Enthalpy_root} can be expressed directly through equation \ref{EQ:Enthalpy_PR} as presented in \ref{CH:Enthalpy} or given by \citet{Gmehling2019}.

@@

		\todo[inline]{Add a short section about boundary conditions and comment on the used numerical methods}

@@

		The efficiency of the process (the yield) is calculated according to equations \ref{Model_measurment_1} to \ref{Model_measurment_2}, which evaluate the mass of solute at the exit of the extraction unit and sums it. The integral form of the measurement equation can be transformed into the differential form and augmented with model equations to be solved simultaneously.

@@

		\subsubsection{State-space representation} \label{CH: State_space}
		It is assumed that the solvent is free of solute at the entrance of the extractor and that all the solid particles have the same initial solute content ${\color{magenta}c_s_0}$. Moreover, it is considered that the initial temperature of the extractor in every place is equal to ${\color{magenta}h_0}$. Therefore, the initial conditions employed in the simulation are:

		{\footnotesize
			\begin{subequations}
				\begin{align*}
					{\color{blue}c_f}(t = 0, z) &= 0   \\
					{\color{blue}c_s}(t = 0, z) &= {\color{magenta}c_s_0} \\
					{\color{blue}h}(t = 0, z) &= {\color{magenta}h_0}
				\end{align*}
		\end{subequations} }

@@

		{\footnotesize
			\begin{align}
				\begin{bmatrix}
					\cfrac{\partial {\color{blue}c_f}(t,z)}{\partial t}\\
					\cfrac{\partial {\color{blue}c_s}(t,z)}{\partial t}\\
					\cfrac{\partial {\color{blue}T}(t,z)}{\partial t} 
				\end{bmatrix}
				& =
				\begin{bmatrix}
					{\color{blue}\phi_1} \left( {\color{blue}c_f}(t,z),{\color{blue}c_s}(t,z),{\color{blue}h}(t,z); {\color{magenta}\Theta} \right)\\
					{\color{blue}\phi_2} \left( {\color{blue}c_f}(t,z),{\color{blue}c_s}(t,z),{\color{blue}h}(t,z); {\color{magenta}\Theta} \right)\\
					{\color{blue}\phi_3} \left( {\color{blue}c_f}(t,z),{\color{blue}c_s}(t,z),{\color{blue}h}(t,z); {\color{magenta}\Theta} \right)
				\end{bmatrix} = {\color{blue}\phi} \left( t,z; {\color{magenta}\Theta} \right) = \cfrac{\partial {\color{blue}\chi}(t,z)}{\partial t}
		\end{align} }

@@

		The method of lines is used to transform the process model equations into a set of ODEs denoted as ${\color{blue}G}({\color{blue}x}(t);{\color{magenta}p})$. The partial derivatives in $z$-direction are computed using a first-order and second-order finite difference approximation. The backward finite difference is used to approximate the first-order derivative, while the central difference scheme is used to approximate the second-order derivative. The length of the fixed bed is divided into $N_z$ equally distributed points in $z$-direction. Each function ${\color{blue}\phi_i}$ is transformed to a corresponding set of $N_z$ discretized equations denoted as ${\color{blue}G}_{i\times N_z+1}$ to ${\color{blue}G}_{(i+1)\times N_z}$, where $i$ corresponds to the process model equation. The state-space model ${\color{blue}\chi}(t,z)$ after the discretization is represented by $\dot{{\color{blue}x}}(t)$ (equation  \ref{discretization}).

			{\footnotesize
				\begin{align*} \label{discretization}
					\dot{{\color{blue}x}}(t) &= \cfrac{d {\color{blue}x}(t)}{d t} = 
					\begin{bmatrix}
						\cfrac{d {\color{blue}c_{f,1}}(t)}{d t} 	  \\
						\vdots					  \\
						\cfrac{d {\color{blue}c_{f,N_z}}(t)}{d t} \\
						\\ \hline \\
						\cfrac{d {\color{blue}c_{s,1}}(t)}{d t} 	  \\
						\vdots					  \\
						\cfrac{d {\color{blue}c_{s,N_z}}(t)}{d t} \\
						\\ \hline \\
						\cfrac{d {\color{blue}h_1}(t)}{d t} 	  \\
						\vdots 					  \\
						\cfrac{d {\color{blue}h_{N_z}}(t)}{d t}
					\end{bmatrix}
					=
					\underbrace{\begin{bmatrix}
							{\color{blue}G_1} \left( {\color{blue}c_f}(t),{\color{blue}c_s}(t),{\color{blue}h}(t); {\color{magenta}p} \right)\\ 
							\vdots\\ 
							{\color{blue}G_{N_z}} \left( {\color{blue}c_f}(t),{\color{blue}c_s}(t),{\color{blue}h}(t); {\color{magenta}p} \right)\\ 
							\\ \hline \\ \\
							{\color{blue}G_{N_z+1}} \left( {\color{blue}c_f}(t),{\color{blue}c_s}(t),{\color{blue}T}(t); {\color{magenta}p} \right)\\ 
							\vdots\\
							{\color{blue}G_{2N_z}} \left( {\color{blue}c}(t),{\color{blue}q}(t),{\color{blue}T}(t); {\color{magenta}p} \right)\\ 
							\\ \\ \hline \\ 
							{\color{blue}G_{2N_z+1}} \left( {\color{blue}c_f}(t),{\color{blue}c_s}(t),{\color{blue}h}(t); {\color{magenta}p} \right) \\
							\vdots\\
							{\color{blue}G_{3N_z}} \left( {\color{blue}c_f}(t),{\color{blue}c_s}(t),{\color{blue}h}(t); {\color{magenta}p} \right)
					\end{bmatrix}}_{{\color{blue}G} \left( {\color{blue}x}(t); {\color{magenta}p} \right)} 
			\end{align*} }

			where ${\color{blue}x} \in \mathbb{R}^{N_x = 3N_z} $ and ${\color{magenta}p} \in \mathbb{R}^{N_p =  N_{\Theta} + N_u } $, $N_{\Theta}$ is the number of model parameters, $N_{u}$ is the number of control variables.

@@

### 2023-04-22 SFE-article/Sections/Parameter_estimation.tex
21.04

The goal of parameter estimation is to obtain the "best" estimate of parameter set $\theta$ (which is a subset of the parameter space $\Theta$ containing all parameters of a model) based on the continuous observations $Y(t)$ or the discrete observations $Y(t_i)$. Conceptually, the unobservable error $\epsilon(t)$ is added to the deterministic model output, $y(t)$ (equation \ref{Model_measurment_1}), to give the observable dependent variable $Y(t)$ (for example results of an experiments). For discrete observations, this can be expressed as:

@@

In the case where the error in the first observation is denoted as $\epsilon_1$, the error in the second observation $\epsilon_2'$ incorporates $\epsilon_1$ as well as an independent random component, given by $\epsilon_2' = \epsilon_1 + \epsilon_2$. Similarly, the error in the third observation is $\epsilon_3' = \epsilon_1 + \epsilon_2 + \epsilon_3$, and so on. \citet{Mandel1957}  made a distinction between the typically assumed independent measurement error in the dependent variable and a "cumulative" or interval error, in which each new observation encompasses the error of the previous ones. Cumulative errors arise from fluctuations in the process itself due to small variations in operating conditions and are not independent; only the differences in measurement from one period to the next are independent.

\subsubsection{Least squares estimation}
In statistics, least squares parameter estimation is a method used to estimate the parameters of a linear regression model. One of the benefits of this method is that it does not require any prior knowledge about the distribution of unobservable errors. Additionally, it yields unbiased estimates and results in the minimum variance among all linear unbiased estimators.

If the model responses, represented by $Y$, are continuous functions of time from $t = 0$ to $t = t_f$, the Markov criterion, also known as "rigorous least squares," is used to minimize the integral squared error $\gamma$. This is achieved by minimizing the following expression:

@@

One characteristic feature of certain special experiments is the use of the same batch of materials for the entire series of measurements. If unobservable error in the first observation is designated as $\epsilon_1$, the second error observation $\epsilon_2'$ includes $\epsilon_1$ plus a random component introduced side from $\epsilon_1$, or $\epsilon_2' = \epsilon_1 + \epsilon_2$. The error in the third observation is $\epsilon_3' = \epsilon_1 + \epsilon_2 + \epsilon_3$, and so forth. \citet{Mandel1957} distinguished between the usually assumed type of independent measuring error in the dependent variable and a 'cumulative' or interval error in which each new observation includes the error of the previous observations. Cumulative errors, arising because of the fluctuations as a function of time in the process itself due to small changes in operating, are not independent - only the differences in measurement from one period to the next are independent. Thus, if we consider two model:

@@

Maximum likelihood estimation (MLE) is a method of estimating the parameters of an assumed probability distribution, given some observed data. This is achieved by maximizing a likelihood function so that, under the assumed statistical model, the observed data is most probable. The MLE has the desirable characteristics of asymptotic efficiency and normality. Each time it has been associated with the (joint) normal distribution because of mathematical convenience. Consider the joint probability density function (the likelihood function) $p \left(\theta | y(t_1),y(t_2),...,y(t_n)\right)$ for parameter $\theta$. If a maximum of this function over all choices of $\theta$ can be found, the estimates so obtained are maximum likelihood estimates. The conditions at the maximum can be evolved incorporating prior information as follows.

The posterior probability density $p(\theta|y)$ can be expressed as the ratio of two probability densities if we make use of the analogue for continuous variables of Equation \ref{EQ: Probabiltiy_independet_b}:

@@

We can get rid of the evidence term $p\left( y(t_n),...,y(t_1) \right)$ because it's constant with respect to the maximization. Moreover, if we are lacking a prior distribution over the quantity we want to estimate, then $p(\theta)$ can be omitted. In such a case:

@@

By collecting a values of $y$ and selecting the values of $\theta$ that maximize the likelihood function $L\left( \theta | y \right)$, a function described in chapter \ref{CH: Bayes} in connection with Bayes' theorem\todo{Is the reference it previous chapter needed?}. Such estimators, $\hat{\theta}$, are known as maximum likelihood estimators. In effect, the methods selects those values of $\theta$ that are at least as likely to generate the observed sample as any other set of values of the parameters if the probability density of $y$ were to be extensively simulated through use of the probability density $p\left( y|\theta \right)$. In making a maximum likelihood estimate, we assume the form of the probability density and that all possible values of $\theta$ are equally likely before experimentation. \todo{Needed?}

@@

		Based on the first order optimality condition, the $\ln L$ can be maximized with respect to the vector $\theta$ by equating to zero the partial derivatives of $\ln L$ with respect to each of the parameters:

@@

		It can be shown that as $n$ approaches infinity the maximum likelihood estimates have the desirable asymptotic properties:

@@

		Maximum likelihood estimates are not necessarily unbiased. Maximum likelihood estimates, however are efficient and, hence, consistent estimates. Furthermore, where a sufficient estimate can be obtained, the maximum likelihood method will obtain it. Finally, if $\hat{\theta}$ is a maximum likelihood estimator of $\theta$, the $f\left( \hat{\theta} \right)$ is a maximum likelihood estimator of $f\left( \theta \right)$, a function of $\theta$.

@@

		%Under the assumption of normality, the lack of prior knowledge about the parameters $\theta^0$ and known initial conditions $y_0$. Intuitively, this selects the parameter values that make the observed data most probable. The specific value $\hat{\theta}$ that maximizes the likelihood function $L$ is called the maximum likelihood estimate.

		As already discussed, the prior information about the estimated parameters can be incorporated in to the maximum likelihood estimator to form a maximum a posteriori probability (MAP) estimate. MAP can be used to obtain a point estimate of an unobserved quantity on the basis of empirical data. MAP estimation can therefore be seen as a regularization of maximum likelihood estimation. The relationship between MAP and MLE can be obtained from the Bayes' rule:

>>>>

The goal of parameter estimation is to obtain the "best" estimate of parameter set $\theta$ (which is a subset of the parameter space $\Theta$ containing all parameters of a model) based on the continuous observations $Y(t)$ or the discrete observations $Y(t_i)$. Conceptually, the unobservable error $\epsilon(t)$ is added to the deterministic model output, $y(t)$ (equation \ref{Model_measurment_1}), to give the observable dependent variable $Y(t)$ (for example results of experiments). For discrete observations, this can be expressed as:

@@

When the error in the first observation is denoted as $\epsilon_1$, the error in the second observation $\epsilon_2'$ incorporates $\epsilon_1$ and an independent random component, given by $\epsilon_2' = \epsilon_1 + \epsilon_2$. Similarly, the error in the third observation is $\epsilon_3' = \epsilon_1 + \epsilon_2 + \epsilon_3$, and so on. \citet{Mandel1957}  distinguished between the typically assumed independent measurement error in the dependent variable and a "cumulative" or interval error, in which each new observation encompasses the error of the previous ones. Cumulative errors arise from fluctuations in the process due to small variations in operating conditions and are not independent; only the differences in measurement from one period to the next are independent.

\subsubsection{Least squares estimation}
In statistics, least squares parameter estimation is a method used to estimate the parameters of a linear regression model. One of the benefits of this method is that it does not require prior knowledge about the distribution of unobservable errors. Additionally, it yields unbiased estimates and results in the minimum variance among all linear unbiased estimators.

If the model responses, represented by $Y$, are continuous time functions from $t = 0$ to $t = t_f$, the Markov criterion, also known as "rigorous least squares," is used to minimize the integral squared error $\gamma$. This is achieved by minimizing the following expression:

@@

One characteristic feature of certain special experiments is the usage of the same batch of materials for the entire series of measurements. If unobservable error in the first observation is designated as $\epsilon_1$, the second error observation $\epsilon_2'$ includes $\epsilon_1$ plus a random component introduced side from $\epsilon_1$, or $\epsilon_2' = \epsilon_1 + \epsilon_2$. The error in the third observation is $\epsilon_3' = \epsilon_1 + \epsilon_2 + \epsilon_3$, and so forth. \citet{Mandel1957} distinguished between the usually assumed type of independent measuring error in the dependent variable and a 'cumulative' or interval error in which each new observation includes the error of the previous observations. Cumulative errors, arising because of the fluctuations as a function of time in the process due to small operating changes, are not independent - only the differences in measurement from one period to the next are independent. Thus, if we consider two models:

@@

Maximum likelihood estimation (MLE) is a method of estimating the parameters of an assumed probability distribution given some observed data. This is achieved by maximizing a likelihood function so that the observed data is most probable under the assumed statistical model. The MLE has the desirable characteristics of asymptotic efficiency and normality. Each time it has been associated with the (joint) normal distribution because of mathematical convenience. Consider the joint probability density function (the likelihood function) $p \left(\theta | y(t_1),y(t_2),...,y(t_n)\right)$ for parameter $\theta$. If a maximum of this function over all choices of $\theta$ can be found, the estimates so obtained are maximum likelihood estimates. The conditions at the maximum can be evolved incorporating prior information as follows.

The posterior probability density $p(\theta|y)$ can be expressed as the ratio of two probability densities if we make use of the analog for continuous variables of Equation \ref{EQ: Probabiltiy_independet_b}:

@@

We can get rid of the evidence term $p\left( y(t_n),...,y(t_1) \right)$ because it's constant with respect to the maximization. Moreover, if we lack a prior distribution over the quantity we want to estimate, then $p(\theta)$ can be omitted. In such a case:

@@

By collecting values of $y$ and selecting the values of $\theta$ that maximize the likelihood function $L\left( \theta | y \right)$, a function described in chapter \ref{CH: Bayes} in connection with Bayes' theorem\todo{Is the reference it previous chapter needed?}. Such estimators, $\hat{\theta}$, are known as maximum likelihood estimators. In effect, the methods select those values of $\theta$ that are at least as likely to generate the observed sample as any other set of values of the parameters if the probability density of $y$ were to be extensively simulated through the use of the probability density $p\left( y|\theta \right)$. In making a maximum likelihood estimate, we assume the form of the probability density and that all possible values of $\theta$ are equally likely before experimentation. \todo{Needed?}

@@

		Based on the first-order optimality condition, the $\ln L$ can be maximized with respect to the vector $\theta$ by equating to zero the partial derivatives of $\ln L$ with respect to each of the parameters:

@@

		It can be shown that as $n$ approaches infinity, the maximum likelihood estimates have the desirable asymptotic properties:

@@

		Maximum likelihood estimates are not necessarily unbiased. Maximum likelihood estimates, however, are efficient and consistent estimates. Furthermore, where a sufficient estimate can be obtained, the maximum likelihood method will obtain it. Finally, if $\hat{\theta}$ is a maximum likelihood estimator of $\theta$, the $f\left( \hat{\theta} \right)$ is a maximum likelihood estimator of $f\left( \theta \right)$, a function of $\theta$.

@@

		%Under the assumption of normality, the lack of prior knowledge about the parameters $\theta^0$ and known initial conditions $y_0$. Intuitively, this selects the parameter values that make the observed data most probable. The maximum likelihood estimate is the specific value $\hat{\theta}$ that maximizes the likelihood function $L$.

		As already discussed, the prior information about the estimated parameters can be incorporated in to the maximum likelihood estimator to form a maximum a posteriori probability (MAP) estimate. MAP can be used to obtain a point estimate of an unobserved quantity based on empirical data. MAP estimation can therefore be seen as a regularization of maximum likelihood estimation. The relationship between MAP and MLE can be obtained from the Bayes' rule:

### 2023-04-22 SFE-article/Sections/Porosity.tex
21.04

    We can first calculate the volume excluding the pore space using the true density of the material:

@@

    We can then use the volume of the solid particles including the pore space and the volume excluding the pore space to calculate the porosity:

@@

    Therefore, the porosity of the solid particles is 0.5, or 50\%. This means that half of the total volume of the solid particles is made up of pore space.

>>>>

    We can first calculate the volume, excluding the pore space, using the true density of the material:

@@

    We can then use the volume of the solid particles, including the pore space and the volume excluding the pore space, to calculate the porosity:

@@

    Therefore, the porosity of the solid particles is 0.5 or 50\%. This means that half of the total volume of the solid particles is made up of pore space.
